---
title: "Raster data downloads and calculations"
author: "Natalie Love"
date: "2/6/2020"
output: html_document
---

##Load packages
```{r}
library(tidyverse)
library(raster)
library(rgdal)
library(sf)
```

##Load data
```{r}
##Load data
nemodata <- read_csv("nemo_occurrences.csv", col_types = cols(
  id = col_double(),
  institutionCode = col_character(),
  collectionCode = col_character(),
  collectionID = col_character(),
  basisOfRecord = col_character(),
  ownerInstitutionCode = col_character(),
  basisOfRecord = col_character(),
  occurrenceID = col_character(),
  catalogNumber = col_character(),
  otherCatalogNumbers = col_character(),
  kingdom = col_character(),
  phylum = col_character(),
  class = col_character(),
  order = col_character(),
  family = col_character(),
  scientificName = col_character(),
  taxonID = col_double(),
  scientificNameAuthorship = col_character(),
  genus = col_character(),
  specificEpithet = col_character(),
  taxonRank = col_character(),
  infraspecificEpithet = col_character(),
  identifiedBy = col_character(),
  dateIdentified = col_character(),
  identificationReferences = col_character(),
  identificationRemarks = col_character(),
  taxonRemarks = col_character(),
  identificationQualifier = col_character(),
  typeStatus = col_character(),
  recordedBy = col_character(),
  recordNumber = col_character(),
  eventDate = col_character(),
  year = col_double(),
  month = col_double(),
  day = col_double(),
  startDayOfYear = col_double(),
  endDayOfYear = col_double(),
  verbatimEventDate = col_character(),
  occurrenceRemarks = col_character(),
  habitat = col_character(),
  fieldNumber = col_character(),
  informationWithheld = col_character(),
  dataGeneralizations = col_character(),
  dynamicProperties = col_character(),
  associatedTaxa = col_character(),
  reproductiveCondition = col_character(),
  establishmentMeans = col_character(),
  lifeStage = col_character(),
  sex = col_character(),
  individualCount = col_double(),
  preparations = col_character(),
  country = col_character(),
  stateProvince = col_character(),
  county = col_character(),
  municipality = col_character(),
  locality = col_character(),
  locationRemarks = col_character(),
  decimalLatitude = col_double(),
  decimalLongitude = col_double(),
  geodeticDatum = col_character(),
  coordinateUncertaintyInMeters = col_double(),
  verbatimCoordinates = col_character(),
  georeferencedBy = col_character(),
  georeferenceProtocol = col_character(),
  georeferenceSources = col_character(),
  georeferenceVerificationStatus = col_character(),
  georeferenceRemarks = col_character(),
  minimumElevationInMeters = col_double(),
  maximumElevationInMeters = col_double(),
  minimumDepthInMeters = col_double(),
  maximumDepthInMeters = col_double(),
  verbatimDepth = col_character(),
  verbatimElevation = col_character(),
  disposition = col_character(),
  language = col_character(),
  recordEnteredBy = col_character(),
  modified = col_character(),
  rights = col_character(),
  rightsHolder = col_character(),
  accessRights = col_character(),
  recordId = col_character(),
  references = col_character()
))
View(nemodata)
```

##Load cliamte data reference sheets
These will be used to tell R which climate rasters to average or add together to calculate seasonal variables
```{r}
winterppt_vars <- read_csv("winterppt_vars.csv")
springtmean_vars <- read_csv("springtmean_vars.csv")
```



##Function to create sums or averages
```{r}
winter_sum <- function(x,y,z){
  return(x + y + z)
}

spring_mean <- function(x,y,z){
  return((x + y + z)/3)
}
```

##set path
Path for output files
```{r}
path_ppt = "prism_seasonal/ppt/"
path_tmean ="prism_seasonal/tmean/"
```

##Loop to calculate winter PPT for each year
```{r}
for (i in 1896:2018){
  var1 <- raster(winterppt_vars$var1[which(winterppt_vars$name == i)])
  var2 <-raster(winterppt_vars$var2[which(winterppt_vars$name == i)])
  var3 <-raster(winterppt_vars$var3[which(winterppt_vars$name == i)])
  ppt_raster <- overlay(var1, var2, var3, fun = winter_sum)
  writeRaster(x = ppt_raster , filename = file.path(path_ppt, paste0("winterppt_",i)))
}
```

##Loop to calculate spring Tmean for each year
```{r}
for (i in 1896:2018){
  var1 <- raster(springtmean_vars$var1[which(springtmean_vars$name == i)])
  var2 <-raster(springtmean_vars$var2[which(springtmean_vars$name == i)])
  var3 <-raster(springtmean_vars$var3[which(springtmean_vars$name == i)])
  tmean_raster <- overlay(var1, var2, var3, fun = spring_mean)
  writeRaster(x = tmean_raster, filename = file.path(path_tmean, paste0("springtmean_",i)))
}
```


#Bring in raster, test again
Get point specific values to compare to PRISM explorer data
```{r}
tmean2013_test <- raster("prism_seasonal/tmean/springtmean_2013")

nemotmean <- enframe(x = raster::extract(tmean2013_test, nemodata[,c("decimalLongitude","decimalLatitude")]), value = "springtmean_2013", name = NULL)
nemotmean

```

